---
title: "Lab 6"
author: "Ryan Fernald"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, **follow the instructions below and use R Markdown to create a pdf document with your code and answers to the following questions on Gradescope.** You may find a template file by clicking "Code" in the top right corner of this page.

## A. Basic functions

Use the following code to create a list of four matrices:

```{r, echo = T, eval = T}
set.seed(100)
matrix_list <- list(
  A = diag(5),
  B = matrix(rnorm(9), nrow = 3, ncol = 3),
  C = matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2),
  D = diag(c(1:5))
)
matrix_list
```

1.  Use the `lapply` function to create a list of length four containing the inverse of these four matrices.

    ```{r}

    new_matrix <- lapply(matrix_list, solve)
    new_matrix
    ```

2.  Use the `sapply` function to create a vector of length four containing the determinants of these four matrices.

    ```{r}
    another_matrix <- sapply(matrix_list, det)
    another_matrix
    ```

## B. Skewness and Kurtosis

Skewness describes how asymmetrical the distribution of a numerical variable is around its mean. Given observations $x_1,\ldots, x_n$, we can calculate the sample skewness $s$ of a variable using the following formula:

$$s = \frac{\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^3}{\left[\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^2\right]^{3/2}}$$ Kurtosis is a measure of the "tailedness" of the distribution of a numerical variable is around its mean. Higher values of kurtosis indicate more extreme outliers. Given observations $x_1,\ldots, x_n$, we can calculate the sample kurtosis $k$ of a variable using the following formula:

$$k = \frac{\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^4}{\left[\frac{1}{n}\sum\limits_{i=1}^n(x_i-\overline{x})^2\right]^{2}}-3$$

3.  Write a function `skewness()` that takes as input a numeric vector `x` and returns the sample skewness. There are functions in R that compute skewness, but you cannot use any of them--write your own implementation. You may remove all `NA` values by default. Use your function to compute the sample skewness of the `arr_delay` variable in the `flights` dataset contained in the `nycflights13` package.

```{r}
library(tidyverse)
library(nycflights13)
library(palmerpenguins)
```

```{r}
skewness <- function(x, na.rm = TRUE) {
  if (na.rm) {
    x <- na.omit(x)
  }
  n <- length(x)
  mean_x <- mean(x)
  sd_x <- sd(x)
  skewness <- sum((x - mean_x) ** 3) / (n * sd_x **3)
  return(skewness)
}
  
```

```{r}

skew <- skewness(flights$arr_delay)

cat("skewness:", (skew))
```

4.  Write a function `kurtosis()` that takes as input a numeric vector `x` and returns the sample skewness. There are functions in R that compute kurtosis, but you cannot use any of them--write your own implementation. You may remove all `NA` values by default. Use your function to compute the sample kurtosis of the `arr_delay` variable in the `flights` dataset contained in the `nycflights13` package.

```{r}
kurtosis <- function(x, na.rm = TRUE) {
  if (na.rm) {
    x <- na.omit(x)
  }
  n <- length(x)
  mean_x <- mean(x)
  sd_x <- sd(x)
  kurtosis <- sum((x - mean_x) ** 4) / (n * sd_x ** 4) - 3
  return(kurtosis)
}
```

```{r}
kurt <- kurtosis(flights$arr_delay)

cat("kurtosis:", kurt)
```

5.  Write a function `get_column_skewness()` that takes as input a data frame and calculates the skewness of each **numeric** variable. The output should be a data frame with two variables: `variable` containing the name of the variable and `skewness` containing the skewness. Your output data frame should only include the numeric variables. You may remove all `NA` values by default. Demonstrate your function on the `penguins` dataset.

```{r}
get_column_skewness <- function(data, na.rm = TRUE) {
  numeric_columns <- sapply(data, is.numeric) #filter only numeric columns
  data_numeric <- data[, numeric_columns]  #specify row (all rows) and column (only the numeric ones)
  
  skewness_values <- sapply(data_numeric, skewness, na.rm = na.rm)
  
  skewness_df <- data.frame(variable = names(skewness_values), skewness = skewness_values)
  
  return(skewness_df)
}
```

```{r}
data(penguins)

skewness_penguins <- get_column_skewness(penguins)

print(skewness_penguins)
```

## C. Finding an error

Suppose you have two teams of runners participating in a 5k. We wish to write a function that takes as input two vectors representing the times of the runners in each team and returns a list of two vectors representing the ranks of each team's runners.

For example, if the first team's times are `c(16.8, 21.2, 19.1)` and the second team's times are `c(17.2, 18.1, 20.0)`, the function should return `c(1, 6, 4)` for the first team and `c(2, 3, 5)` for the second team.

Below is a draft version of the function `get_runner_ranks()`. However, there is an error somewhere. Use any method we discussed in class to identify the error.

```{r, error = T, echo = T, eval = T}
get_runner_ranks1 <- function(x, y) {
  # combine all runner times
  combined_times <- c(x, y) 
  
  # sort all runner times from fastest to slowest
  sort(combined_times, decreasing = FALSE)
  
  # create ranks vectors
  ranks_x <- numeric(length(x))
  ranks_y <- numeric(length(y))
  
  for (i in seq_along(ranks_x)) {
    # look up rank of time i in x in combined_times
    ranks_x[i] <- match(x[i], combined_times)
  }
  
  for (i in seq_along(ranks_y)) {
    # look up rank of time i in y in combined_times
    ranks_y[i] <- match(y[i], combined_times)
  }
  
  # return a list of first team and second team ranks
  return(list(x = ranks_x, y = ranks_y))
}
```

```{r}
get_runner_ranks1(c(20000000000.8, 51.2, 19123.1), c(171234.2, 132.1, 20.0))
```

6.  Explain in your own words what the error was.

    When I check placement of each runner in this function it seems like the first time recorded is considered first. Because the display is shown as 1, 2, 3, 4, 5, 6. Regardless of what changes I make to the times when I test, which tells me there is a problem in sorting. When we call the sort function it should be stored in a variable so we can use it later.

7.  Below, write a corrected version of `get_runner_ranks()` and compute `get_runner_ranks(c(16.8, 21.2, 19.1), c(17.2, 18.1, 20.0))`.

```{r, error = T, echo = T, eval = T}
get_runner_ranks <- function(x, y) {
  # combine all runner times
  combined_times <- c(x, y) 
  
  # sort all runner times from fastest to slowest
  combined_times <- sort(combined_times, decreasing = FALSE)
  
  # create ranks vectors
  ranks_x <- numeric(length(x))
  ranks_y <- numeric(length(y))
  
  for (i in seq_along(ranks_x)) {
    # look up rank of time i in x in combined_times
    ranks_x[i] <- match(x[i], combined_times)
  }
  
  for (i in seq_along(ranks_y)) {
    # look up rank of time i in y in combined_times
    ranks_y[i] <- match(y[i], combined_times)
  }
  
  # return a list of first team and second team ranks
  return(list(x = ranks_x, y = ranks_y))
}
```

```{r}
get_runner_ranks(c(16.8, 21.2, 19.1), c(17.2, 18.1, 20.0))
```
