---
title: "Check-in 1: R Markdown"
author: "Ryan Fernald"
date: "2024-01-29"
output: html_document
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

1.  This is a code chunk that opens the `mtcars` dataset, which contains information about on 32 cars from a 1974 issue of the magazine *Motor Trend* and creates a scatter plot with horsepower `hp` on the x-axis and miles-per-gallon `mpg` on the y-axis. Try running this code and Rendering/Knitting your Rmd document. If you have done this correctly, RStudio should produce an html document.

```{r, eval = T}
data(mtcars)
plot(mtcars$hp, mtcars$mpg)
```

2.  Using this scatter plot, describe the relationship between `mpg` and `hp`.

The data appears to have a slight negative correlation between, meaning the lower the horse power, the higher the MPG, conversely the higher the horse power the lower the MPG.

3.  Create your own code chunk below this question containing the line of code `print("Hello MATH167R!")`. Then Render/Knit your Rmd document. If you have done this correctly, RStudio should produce an html document.

```{r, eval = T}
print("Hello MATH167R!")
```

$$
 \begin{array}{l}I_{n}( x) =\frac{2n}{x}\int _{-1}^{1} z\left( 1-z^{2}\right)^{n-1} sin( xz) \ dz=\frac{2n}{x^{2}}\int _{-1}^{1}\left[\left( 1-z^{2}\right)^{n-1} -2z^{2}( n-1)\left( 1-z^{2}\right)^{n-2}\right] \cdotp cos( xz) \ dz\\\\\cdots \\\\I_{n}( x) =\frac{2n}{x}\int _{-1}^{1} z\left( 1-z^{2}\right)^{n-1} sin( x\cdotp \ z) \ dz\\\\\begin{array}{|l|c|}\hlineu=z\cdotp \left( 1-z^{2}\right)^{n-1} & dv=sin( xz) \ dz\\\hlinedu=\left( 1-z^{2}\right)^{n-1} +z\cdotp \left(( n-1)\left( 1-z^{2}\right)^{n-2} \cdotp ( -2z)\right) \ dz & v=-\frac{1}{x} \cdotp cos( xz)\\\hline\end{array}\\\\\int udv=uv-\int vdu\\\\{\textstyle =\frac{2n}{x}\left[\left[ z\cdotp \left( 1-z^{2}\right)^{n-1}\frac{1}{x} \cdotp cos( xz)\right]_{-1}^{1} -\int _{-1}^{1} -\frac{1}{x} \cdotp cos( xz) \cdotp \left(\left( 1-z^{2}\right)^{n-1} +z\cdotp ( n-1)\left( 1-z^{2}\right)^{n-2} \cdotp ( -2z)\right) \ dz\right]}\\\\=0+\frac{2n}{x^{2}}\int _{-1}^{1}\left[\left( 1-z^{2}\right)^{n-1} -2z^{2}( n-1)\left( 1-z^{2}\right)\right] \cdotp cos( xz) \ dz\\\\=\frac{2n}{x^{2}}\int _{-1}^{1}\left[\left( 1-z^{2}\right)^{n-1} -2z^{2}( n-1)\left( 1-z^{2}\right)^{n-2} \cdotp cos( xz) \ dz\right]\end{array}
$$
