---
title: "Check-in 2"
author: "Ryan Fernald"
date: ""
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, **follow the instructions below and use R Markdown to create a pdf document with your code and answers to the following questions on Gradescope.** You may find a template file by clicking "Code" in the top right corner of this page.

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

## A. Flights Data

**1.** For this lab, we'll be using the `flights` data. You can load this data using the following code:

```{r}
library(nycflights13)
data(flights)
```

**2.** Use the `mutate()` function to turn `origin` into a factor.

```{r}
library(tidyverse)
flights <- flights |>
  mutate(origin = as.factor(origin))

class(flights$origin)
```

**3.** Compute the mean `arr_delay` for each origin airport. Which airport has the longest delays on average? You can drop missing rows.

```{r}
flights |>
  group_by(origin) |>
  summarize(flights_arr_delay = mean(arr_delay, na.rm = T)) |>
  ungroup()
```

The airport with the longest delays on average is EWR, Newark Airport.

By using the `na.rm = T`, we are essentially dropping the missing rows.

**4.** Use the `mutate()` function to turn `month` into a factor.

```{r}
flights <- flights |>
  mutate(month = as.factor(month))
class(flights$month)
```

**5.** Compute the mean `arr_delay` by month. What do you notice?

```{r}
flights |>
  group_by(month) |>
  summarize(month_arr_delay = mean(arr_delay, na.rm = T))
```

The output is a tibble with a 12 x 2 dimensions. First thing I noticed was the month is a factor instead of an integer, and the month_arr_delay is a double. In questions 4 we changed the month column into a column of factors instead of integers.

**6.** Compute the mean `arr_delay` by month AND airport. What do you notice?

```{r}
flights |>
  group_by(month,origin) |>
  summarize(month_arr_delay = mean(arr_delay, na.rm = T))

```

By using group_by() we can filter the data by specified from multiple columns of data. From there we can preform what ever calculation we are interested in getting and creating a nice tibble to display our filtered data. This tibble shows us each month broken down with each airport's monthly average delay.
